---
title: "Dynamic simulation of forest growth over 300 years with global sensitivity analysis"
author: "Fletcher McConnell & Linus Ghanadan"
date: "05-16-2024"
output: html_document
---

```{r setup, message=FALSE}
# load packages
library(tidyverse)
library(deSolve)
library(sensitivity)
```

## Setup

```{r}
# load packages
library(tidyverse)
library(deSolve)
library(sensitivity)

# source forest growth function
source("dforestgrowth.R")
```

## Run model for 300 years (using the ODE solver)

```{r}
# set number of parameters
np = 2000

# generate parameter sets for sensitivity analysis
r_exp = rnorm(mean = 0.01, sd = 0.002, n = np)
g_linear = rnorm(mean = 2, sd = 0.5, n = np)
K_capacity = rnorm(mean = 250, sd = 50, n = np)
threshold_closure = rnorm(mean = 50, sd = 10, n = np)

X1 = cbind.data.frame(r = r_exp, g = g_linear, K = K_capacity, threshold = threshold_closure)

r_exp = rnorm(mean = 0.01, sd = 0.002, n = np)
g_linear = rnorm(mean = 2, sd = 0.5, n = np)
K_capacity = rnorm(mean = 250, sd = 50, n = np)
threshold_closure = rnorm(mean = 50, sd = 10, n = np)

X2 = cbind.data.frame(r = r_exp, g = g_linear, K = K_capacity, threshold = threshold_closure)

# fix negative values
X1 = X1 %>% map_df(pmax, 0.0)
X2 = X2 %>% map_df(pmax, 0.0)

# create Sobol object and get parameter sets for running model
sens_P = sobolSalt(model = NULL, X1, X2, nboot = 300)

# add names to parameter sets
colnames(sens_P$X) = c("r", "g", "K", "threshold")

# set initial forest size and simulation times
Cinitial = 10
simtimes = seq(from = 1, to = 300)

# run model with first parameter set
parms = list(r = sens_P$X[1, "r"], g = sens_P$X[1, "g"], K = sens_P$X[1, "K"], threshold = sens_P$X[1, "threshold"])

result = ode(y = Cinitial, times = simtimes, func = dforestgrowth, parms = parms)
result = as.data.frame(result)
colnames(result) = c("time", "C")

# plot results
ggplot(result, aes(time, C)) +
  geom_line() +
  labs(x = "Time (years)", y = "Forest Size (kg C)", title = "Forest Growth Trajectory") +
  theme_bw()
```

## Sobol global sensitivity analysis

```{r}
# function to compute maximum forest size
compute_max_forest_size <- function(carbontime) {
  max_size <- max(carbontime$C)
  return(list(max_size = max_size))
}

# set number of parameters
np <- 2000

# generate parameter sets for sensitivity analysis
r_exp <- rnorm(mean = 0.01, sd = 0.001, n = np)
g_linear <- rnorm(mean = 2, sd = 0.2, n = np)
K_capacity <- rnorm(mean = 250, sd = 25, n = np)
threshold_closure <- rnorm(mean = 50, sd = 5, n = np)

X1 <- cbind.data.frame(r = r_exp, g = g_linear, K = K_capacity, threshold = threshold_closure)

r_exp <- rnorm(mean = 0.01, sd = 0.001, n = np)
g_linear <- rnorm(mean = 2, sd = 0.2, n = np)
K_capacity <- rnorm(mean = 250, sd = 25, n = np)
threshold_closure <- rnorm(mean = 50, sd = 5, n = np)

X2 <- cbind.data.frame(r = r_exp, g = g_linear, K = K_capacity, threshold = threshold_closure)

# create Sobol object and get parameter sets for running model
sens_forest <- sobolSalt(model = NULL, X1, X2, nboot = 300)
colnames(sens_forest$X) <- c("r", "g", "K", "threshold")

# wrapper function to run model and compute maximum forest size
p_wrapper <- function(r, g, K, threshold, Cinitial, simtimes, odefunc, metricfunc) {
  parms <- list(r = r, g = g, K = K, threshold = threshold)
  result <- ode(y = Cinitial, times = simtimes, func = odefunc, parms = parms)
  result <- as.data.frame(result)
  colnames(result) <- c("time", "C")
  metrics <- metricfunc(result)
  return(metrics)
}

# set initial forest size and simulation times
Cinitial <- 10
simtimes <- seq(from = 1, to = 300)

# run model for all parameter sets and compute maximum forest size
allresults <- as.data.frame(sens_forest$X) %>%
  pmap(p_wrapper, Cinitial = Cinitial, simtimes = simtimes, odefunc = dforestgrowth, metricfunc = compute_max_forest_size)

# extract maximum forest size results into a data frame
allres <- allresults %>% map_dfr(`[`, "max_size")
colnames(allres) <- "max_size"

# create sensitivity analysis box plot
ggplot(allres, aes(x = "All Parameter Sets", y = max_size)) +
  geom_boxplot(color = "black") +
  labs(x = NULL, y = "Maximum Forest Size (kg C)", title = "Global Sensitivity Analysis: Impact of Varying Parameters on Maximum Forest Size") +
  theme_minimal()
```

```{r}
# perform sensitivity analysis
sens_result <- tell(sens_forest, allres$max_size)

# label rows in the Sobol indices
rownames(sens_result$S) <- c("r", "g", "K", "threshold")
rownames(sens_result$T) <- c("r", "g", "K", "threshold")

# print Sobol indices with labeled rows
print(sens_result)
```


## Conclusion

This sensitivity analysis suggests that the carrying capacity (K) and the pre-canopy closure growth rate (r) are the most influential parameters in determining the maximum forest size estimates. Variations in either of these parameters could lead to significant differences in the maximum size that a forest can reach. Because of this, it is crucial to have accurate estimates of all parameters, but especially K and r, so that predictions of the maximum forest size are accurate. 
